Blame and Coercion: Together Again for the Second Time

### 2019-10-16 paper.tex
hyper-coercion => hypercoercion

% Hyper-coercion

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  \todo[inline]{Rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hyper-coercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
  \todo[inline]{Spell it ``hypercoercion'' without the hyphen. -Jeremy}
\end{abstract}

@@

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bi-simulate the 
corresponding Blame Calculi.

@@

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} coercion in \citet{siek2012interpretations} and prove that 
their revised coercion calculus bi-simulates the \lazyUD{} Blame Calculus.

@@

We present hyper-coercion, a cast representation whose composition is 
structurally recursive, and whose definition is self-standing. What's more, it 
should be at 
least as easy to implement and understand as normal coercions, because there is 
a clear connection between them, as we will show in Section 
\ref{sec:hyper-coercion}.
Our hyper-coercion considers sum types and product types, which are not 
considered in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hyper-coercion. This suggests that 
hyper-coercion might scale better than super-coercion.

We prove formally that the \lazyD{} (resp. \lazyUD{}) hyper-coercion calculus 
bi-simulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is 
the first bi-simulation proof for the \lazyD{} Blame Calculus as far as we know.

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ Blame 
Calculi. 
Section \ref{sec:coercion-calculus} reviews the state-of-art of $Lazy$ Coercion 
Calculi. 
In section \ref{sec:hyper-coercion} we present Hyper-coercion.
Section \ref{sec:conclude} concludes.

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}).
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hyper-coercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hyper-coercions.

@@

As the name suggests, 
hyper-coercion is largely inspired by coercion, especially its subst NC. We 
will highlight the similarity of normal coercion and hyper-coercion, and hope 
this would convince you that hyper-coercion should be as difficult to implement 
as NC.

@@

Three observations on this definition leads to hyper-coercion: 
\begin{enumerate}
	\item The length of normal coercion is at most three.
	\item Projections are always at the beginning when presented.
	\item Injections and failures are always at the end when presented.
\end{enumerate}

\section{Hyper-coercion} \label{sec:hyper-coercion}

\begin{figure}
	Syntax
	\[
	\begin{array}{lclr}
	\stxrule{I}{injectable types (\lazyD)}{P}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOunit \mid
		\POOfun{\TOOdyn}{\TOOdyn} \mid
		\POOprod{\TOOdyn}{\TOOdyn} \mid
		\POOsum{\TOOdyn}{\TOOdyn}
	}
	\stxrule{c}{hyper-coercions}{
		\hyperCoercionI \mid{}
		\hyperCoercionC{h}{m}{t}
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{m}{middles}{
		\POOunit \mid
		\POOfun{c_1}{c_2} \mid
		\POOprod{c_1}{c_2} \mid
		\POOsum{c_1}{c_2}
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\end{array}
	\]

	Hyper-coercion typing \fbox{$ c : T \Longrightarrow T $}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

@@

	\caption{Definition of Hyper-coercion (HC)}
	\label{fig:hyper-coercion}
\end{figure}

This section presents our first contribution, hyper-coercion (HC). 
\figref{fig:hyper-coercion} defines the syntax of HC.
Types, pre-types, and blame labels are as before.

Let $ c $ ranges over hyper-coercions. A hyper-coercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ ranges over heads. A head is either a no-op, or a projection.
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to middles $ m \smile m $ in the natural way.
Let $ t $ ranges over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ ranges over no-op and labels. It is useful when we introduce 
composition of hyper-coercions.

\subsection{\lazyD{} Hyper-coercion}

@@

	Composition of hyper-coercions \fbox{$ c \fatsemi^\ell c = c $}
	\[ 
	\begin{array}{rclclr}

@@

	\fbox{$ cast(T,l,T) = c$}
	\[
	\begin{array}{rclr}
	\funrule{cast(T_1,l,T_2)}{
		id(T_1) \fatsemi^l id(T_2)
	}{}
	\end{array}
	\]
	\caption{\lazyD{} Hyper-coercion}
	\label{fig:HC-D}
\end{figure}

@@

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hyper-coercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[\lazyD\ Hyper-coercions is a Monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

@@

\subsection{\lazyUD{} Hyper-coercion [Jeremy]}

@@

We list a fraction of reduction rules due to space limitation.
When values are constructed, their hyper-coercion parts are filled with outputs 
of $ id $. For instance, when a function value is constructed, its first part 
and last part are initialized to $ id(T_1) $ and $ id(T_2) $ respectively.
When evaluating a cast expression, the current continuation is extended with a 
hyper-coercion constructed by $ cast $. $ ext $ composes the new hyper-coercion 
with the hyper-coercion on the top of the continuation.
When evaluating a compound expression, the machine firstly construct the new 
pre-continuation, then turn it to a continuation by adding an identity 
hyper-coercion at the top. For instance, when evaluating a \texttt{cons} 
expression, the machine firstly construct $ \hckOOconsI{e_2}{\rho}{\kappa} $, 
the new pre-continuation, then call $ cont $, which adds an identity cast to 
form a continuation. 
When a function call happens, the machine firstly cast the operand. If the 
casting succeeds, the machine then evaluate the body in the extended 
environment and the extended continuation. If the casting fails, the machine 
then halts with the blame label.

@@

\section{Correctness Proof of \lazyD{} Hyper-coercions}

@@

With the properties of surely \lazyD{} CR, it is straightforward to proof that 
all machines instantiated with these CRs bi-simulate each other. 
The state bisimilarity ($\eqvS{C}{C}{s}{s}$) is 
derived from the definition of states. We omit subscripts when there is no 
ambiguity. \begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}

@@

\subsection{\lazyD\ Hyper-coercion Respect $ \mathcal{D} $}
\label{sec:LDHC-correct}

@@

	\fbox{$ applyTail(t,v) = r $}
	\[
	\begin{array}{rclr}
	\funrule{applyTail(\bot^l,v)}{\rOOfail{l}}{}
	\funrule{applyTail(\epsilon,v)}{\rOOsucc{v}}{}
	\funrule{applyTail(!,v)}{\rOOsucc{(\hcvOOinj{P}{v})}}{v : P}
	\end{array}
	\]
	\caption{\lazyD\ Hyper-coercion's $ applyCast $}
	\label{hc-applyCast}
\end{figure}

\begin{lemma}[\lazyD{} Hyper-coercion is Surely \lazyD]
	\label{thm:hc-surely-lazyD}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{theorem}[\lazyD{} Hyper-coercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:hc-surely-lazyD} and 
	Theorem~\ref{thm:surely-lazyD-correct}.
\end{proof}

>>>>

% Hypercoercion

@@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  \todo[inline]{Rewrite this abstract based on the comparision table. -Jeremy}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
  \todo[inline]{Spell it ``hypercoercion'' without the hyphen. -Jeremy}
\end{abstract}

@@

\citet{siek2012interpretations} present the first \lazyD{} Blame Calculus and 
revised coercion calculi.
They simplify the coercion calculi by only 
working with normal coercions (i.e. coercions in normal form). 
They conjecture that their revised coercion calculi bisimulate the 
corresponding Blame Calculi.

@@

\citet{siek2015blame} revisit the coercion-based approach. They revise the 
\lazyUD{} coercion in \citet{siek2012interpretations} and prove that 
their revised coercion calculus bisimulates the \lazyUD{} Blame Calculus.

@@

We present hypercoercion, a cast representation whose composition is 
structurally recursive, and whose definition is self-standing. What's more, it 
should be at 
least as easy to implement and understand as normal coercions, because there is 
a clear connection between them, as we will show in Section 
\ref{sec:hypercoercion}.
Our hypercoercion considers sum types and product types, which are not 
considered in all proofs above. Adding each of them requires us to add only one 
new constructor to a component of hypercoercion. This suggests that 
hypercoercion might scale better than super-coercion.

We prove formally that the \lazyD{} (resp. \lazyUD{}) hypercoercion calculus 
bisimulates the \lazyD{} (resp. \lazyUD{}) Blame Calculus. This is 
the first bi-simulation proof for the \lazyD{} Blame Calculus as far as we know.

The structure of this paper is as follows. 
Section \ref{sec:blame-calculus} reviews the state-of-art of $Lazy$ Blame 
Calculi. 
Section \ref{sec:coercion-calculus} reviews the state-of-art of $Lazy$ Coercion 
Calculi. 
In section \ref{sec:hypercoercion} we present hypercoercion.
Section \ref{sec:conclude} concludes.

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}).
\todo[inline]{The figure is too tall. You may need to pull some things
  out of the figure, such as the value typing. -Jeremy}
\todo[inline]{Please fill in the missing reduction rules. -Jeremy}
\todo[inline]{Explain why you are using a CEK machine instead of
  reduction semantics. -Jeremy}
\todo[inline]{Change the order of parameters for applyCast to take
  the value first and the cast second. -Jeremy}
\todo[inline]{The definition of values is too loose for a cast.
  You need to restrict the casts to just the inert ones. -Jeremy }
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hypercoercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hypercoercions.

@@

As the name suggests, 
hypercoercion is largely inspired by coercion, especially its subst NC. We 
will highlight the similarity of normal coercion and hypercoercion, and hope 
this would convince you that hypercoercion should be as difficult to implement 
as NC.

@@

Three observations on this definition leads to hypercoercion: 
\begin{enumerate}
	\item The length of normal coercion is at most three.
	\item Projections are always at the beginning when presented.
	\item Injections and failures are always at the end when presented.
\end{enumerate}

\section{hypercoercion} \label{sec:hypercoercion}

\begin{figure}
	Syntax
	\[
	\begin{array}{lclr}
	\stxrule{I}{injectable types (\lazyD)}{P}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOunit \mid
		\POOfun{\TOOdyn}{\TOOdyn} \mid
		\POOprod{\TOOdyn}{\TOOdyn} \mid
		\POOsum{\TOOdyn}{\TOOdyn}
	}
	\stxrule{c}{hypercoercions}{
		\hyperCoercionI \mid{}
		\hyperCoercionC{h}{m}{t}
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{m}{middles}{
		\POOunit \mid
		\POOfun{c_1}{c_2} \mid
		\POOprod{c_1}{c_2} \mid
		\POOsum{c_1}{c_2}
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\end{array}
	\]

	hypercoercion typing \fbox{$ c : T \Longrightarrow T $}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

@@

	\caption{Definition of hypercoercion (HC)}
	\label{fig:hypercoercion}
\end{figure}

This section presents our first contribution, hypercoercion (HC). 
\figref{fig:hypercoercion} defines the syntax of HC.
Types, pre-types, and blame labels are as before.

Let $ c $ ranges over hypercoercions. A hypercoercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ ranges over heads. A head is either a no-op, or a projection.
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to middles $ m \smile m $ in the natural way.
Let $ t $ ranges over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ ranges over no-op and labels. It is useful when we introduce 
composition of hypercoercions.

\subsection{\lazyD{} hypercoercion}

@@

	Composition of hypercoercions \fbox{$ c \fatsemi^\ell c = c $}
	\[ 
	\begin{array}{rclclr}

@@

	\fbox{$ cast(T,l,T) = c$}
	\[
	\begin{array}{rclr}
	\funrule{cast(T_1,l,T_2)}{
		id(T_1) \fatsemi^l id(T_2)
	}{}
	\end{array}
	\]
	\caption{\lazyD{} hypercoercion}
	\label{fig:HC-D}
\end{figure}

@@

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[\lazyD\ hypercoercions is a Monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

@@

\subsection{\lazyUD{} hypercoercion [Jeremy]}

@@

We list a fraction of reduction rules due to space limitation.
When values are constructed, their hypercoercion parts are filled with outputs 
of $ id $. For instance, when a function value is constructed, its first part 
and last part are initialized to $ id(T_1) $ and $ id(T_2) $ respectively.
When evaluating a cast expression, the current continuation is extended with a 
hypercoercion constructed by $ cast $. $ ext $ composes the new hypercoercion 
with the hypercoercion on the top of the continuation.
When evaluating a compound expression, the machine firstly construct the new 
pre-continuation, then turn it to a continuation by adding an identity 
hypercoercion at the top. For instance, when evaluating a \texttt{cons} 
expression, the machine firstly construct $ \hckOOconsI{e_2}{\rho}{\kappa} $, 
the new pre-continuation, then call $ cont $, which adds an identity cast to 
form a continuation. 
When a function call happens, the machine firstly cast the operand. If the 
casting succeeds, the machine then evaluate the body in the extended 
environment and the extended continuation. If the casting fails, the machine 
then halts with the blame label.

@@

\section{Correctness Proof of \lazyD{} hypercoercions}

@@

With the properties of surely \lazyD{} CR, it is straightforward to proof that 
all machines instantiated with these CRs bisimulate each other. 
The state bisimilarity ($\eqvS{C}{C}{s}{s}$) is 
derived from the definition of states.
\todo[inline]{The term ``bisimilarity'' has a different technical meaning from
  ``bisimulation relation''. I don't think you are using it properly
  in the above sentence. -Jeremy}
We omit subscripts when there is no 
ambiguity.
\begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}

@@

\subsection{\lazyD\ hypercoercion Respect $ \mathcal{D} $}
\label{sec:LDHC-correct}

@@

	\fbox{$ applyTail(t,v) = r $}
	\[
	\begin{array}{rclr}
	\funrule{applyTail(\bot^l,v)}{\rOOfail{l}}{}
	\funrule{applyTail(\epsilon,v)}{\rOOsucc{v}}{}
	\funrule{applyTail(!,v)}{\rOOsucc{(\hcvOOinj{P}{v})}}{v : P}
	\end{array}
	\]
	\caption{\lazyD\ hypercoercion's $ applyCast $}
	\label{hc-applyCast}
\end{figure}

\begin{lemma}[\lazyD{} hypercoercion is Surely \lazyD]
	\label{thm:hc-surely-lazyD}
\end{lemma}
\begin{proof}
	See the supplementary material.
\end{proof}

\begin{theorem}[\lazyD{} hypercoercion Respect $ \mathcal{D} $]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	\judgeDBeval{e}{o} \; \text{if and only if} \; 
	\judgeSeval{H}{e}{o}
	\]
\end{theorem}
\begin{proof}
	Immediately from Lemma~\ref{thm:hc-surely-lazyD} and 
	Theorem~\ref{thm:surely-lazyD-correct}.
\end{proof}
